Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and \Spitzer\ Phase Curve Observations

### 2018-02-14 ms.tex
final

\begin{abstract}
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux ratio is $0.151\pm0.015\%$, $0.446\pm0.038\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside.  We fit the phase variation with a range of climate maps and find that a spherical harmonics model generally provides the best fit. The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from $1880\pm40$ K on the nightside to $2930 \pm 10$ K on the dayside. The dayside spectrum has a significantly higher brightness temperature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The inversion is not present on the nightside. The transmission spectrum is featureless, but consistent with the small features expected from cloud-free models. Using the dayside and nightside emission spectra, we retrieved the atmospheric composition with a 1D self-consistent modeling technique and found the atmosphere is moderately metal-enriched ($\mathrm{[M/H]} =  23^{+29}_{-13}\times$ solar) and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to cooler hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and trace amounts of hydrides/oxides.  We compare the phase curves to 3D general circulation models and find that atmospheric drag with a timescale of $10^4$ s is the best match to the data, suggesting the presence of a magnetic field with strength $\sim1$ Gauss.  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find these more massive and self-luminous objects have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.
\end{abstract}

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/w103b_transmission_spectrum.pdf}
\caption{The transmission spectrum of WASP-103b, corrected for planet nightside emission at 1700 K (blue points) and uncorrected (red points). The dark gray line is the model transmission spectrum from the $\tau_\mathrm{drag4}$ GCM. Atmospheric scale height $H$ is shown on the right y-axis, where $H = 5.5\times10^6$ m (assuming a mean molecular weight of 2.3 atomic mass units, surface gravity $g = 15.9$ m/s$^2$, and a temperature $T = 2410$ K).}
\label{fig:tspec}
\end{figure}

@@

\subsection{Blackbody Fits}
\label{sec:bbfits}
A blackbody is the simplest model for the planet's thermal emission and provides a useful first evaluation of the atmospheric properties. To calculate the best fit blackbody model, we assumed a planet-to-star radius ratio of 0.1146 and used a NextGen stellar spectrum interpolated to an effective temperature of 6110 K \citep{allard12}.  The best-fit blackbodies are shown in Figures \ref{fig:spectra} and \ref{fig:dayside}. The model fits the data fairly well: it is consistent with the data at the $3\,\sigma$ level for 70\% of the orbital phases. However, the \Spitzer\ data have higher brightness temperatures than the WFC3 data at phase 0.5, and lower brightness temperatures at phases $0.8 - 0.9$. These differences suggest the presence of an emission feature on the dayside at \Spitzer\ wavelengths, which transitions to an absorption feature on the nightside, perhaps indicating changes in thermal structure with longitude in the atmosphere.

We also fit indepedent blackbody models to the \HST/WFC3 data and each \Spitzer\ channel. The resulting brightness temperatures are listed in Table\,\ref{table:teffs}. The WFC3 data agree well by a blackbody model at all orbital phases except phase 0.5 (consistent at better than $1.5\,\sigma$). The dayside has higher signal-to-noise than the other orbital phases, thanks to the two secondary eclipse observations from \cite{cartier17}. The more sophisticated grid-based retrieval (described in the next section) provides a better fit to the dayside.

@@

\begin{deluxetable}{LLLLL}
	\tablecolumns{5}
	\tablewidth{0pt}
	\tablecaption{Phase-resolved Brightness Temperatures \label{table:teffs}}
	\tablehead{
		\colhead{Orbital Phase} & \colhead{$T_\mathrm{b}$} & \colhead{$T_\mathrm{b}$} & \colhead{$T_\mathrm{b}$} & \colhead{$\chi^2_\nu$}\tablenotemark{1} \\
	\colhead{\,} & \colhead{WFC3} & \colhead{Ch. 1} & \colhead{Ch. 2} & \colhead{(9 dof)}}
		\startdata
		0.06 - 0.15 & 1883 \pm 41 & 1523 \pm 158 & 1589 \pm 98 & 0.7 \\
		0.15 - 0.25 & 2208 \pm 24 & 2612 \pm 104 & 2299 \pm 78 & 0.9 \\
		0.25 - 0.35 & 2587 \pm 16 & 2926 \pm 95 & 2592 \pm 67 & 1.5 \\
		0.35 - 0.44 & 2831 \pm 11 & 2834 \pm 93 & 2976 \pm 61 & 1.5 \\
		0.44 - 0.56 & 2933 \pm 8 & 2995 \pm 141 & 3154 \pm 48 & 2.8 \\
		0.56 - 0.65 & 2811 \pm 11 & 2631 \pm 96 & 2891 \pm 63 & 2.0 \\
		0.65 - 0.75 & 2572 \pm 15 & 2708 \pm 101 & 2421 \pm 78 & 1.3 \\
		0.75 - 0.85 & 2263 \pm 22 & 1952 \pm 117 & 1939 \pm 83 & 1.3 \\
		0.85 - 0.94 & 1987 \pm 34 & 1594 \pm 150 & 1288 \pm 118 & 0.6 \\
		\enddata
		\tablenotetext{1}{$\chi^2_\nu$ values are for the fits to the WFC3 data only.}
	\end{deluxetable}

@@

\begin{deluxetable*}{llCCCC}
	\tablecolumns{6}
	\tablewidth{0pt}
	\tablecaption{Source Properties \label{table:sources}}
	\tablehead{
	\colhead{\,} & \colhead{Object} & \colhead{$T_\mathrm{eff}$ (K)} & \colhead{log\,$g$ (cgs)} & \colhead{H$_2$O A$_1$} & \colhead{H$_2$O A$_2$}}
		\startdata
		Hot Jupiter & W103b night & 1880 \pm 40 & 3.2 \pm 0.04 & 0.07 \pm 1.8\mathrm{e}-01 & -0.00 \pm 1.6\mathrm{e}-01 \\
		\, & W103b quadrature & 2400 \pm 20 & 3.2 \pm 0.04 & -0.14 \pm 7.8\mathrm{e}-02 & -0.01 \pm 6.7\mathrm{e}-02 \\
		\, & W103b dayside & 2930 \pm 10 & 3.2 \pm 0.04 & 0.04 \pm 1.4\mathrm{e}-02 & 0.15 \pm 1.2\mathrm{e}-02 \\
		Brown Dwarf & 2MASS J1320+0409 & 1880 \pm 70 & 5.19 \pm 0.16 & 0.21 \pm 6.3\mathrm{e}-04 & -0.06 \pm 5.0\mathrm{e}-04 \\
		\, & 2MASS J0428-2253 & 2430 \pm 80 & 5.22 \pm 0.09 & 0.16 \pm 1.2\mathrm{e}-04 & -0.03 \pm 1.0\mathrm{e}-04 \\
		\, & 2MASS J0003-2822 & 2890 \pm 80 & 5.18 \pm 0.04 & 0.26 \pm 1.3\mathrm{e}-04 & 0.10 \pm 1.1\mathrm{e}-04 \\
		Imaged Companion & CD-35 2722 & 1800 \pm 100 & 4.5 \pm 0.5 & - & 0.15 \pm 1.0\mathrm{e}-05 \\
		\, & USco 1610-1913B & 2400 \pm 150 & - & 0.27 \pm 2.0\mathrm{e}-04 & 0.19 \pm 2.2\mathrm{e}-04 \\
		\, & TWA 22A & 3000 \pm 100 & 4.5 \pm 0.5 & - & 0.29 \pm 2.0\mathrm{e}-05 \\
		\enddata
	\end{deluxetable*}

@@

	\item{The dayside planet-to-star flux is $0.151\pm0.015\%$, $0.446\pm0.38\%$, and $0.569\pm0.014\%$ in the WFC3 bandpass, \Spitzer\ $3.6$, and \Spitzer\ $4.5\mu$m, respectively.  The best fit blackbody to the WFC3 dayside spectrum has a brightness temperature of $2930 \pm 10$ K, making WASP-103b among the hottest exoplanets ever observed.}

@@

\item{We calculated phase-resolved spectra in ten orbital phase bins. The \HST/WFC3 spectra are consistent with  blackbody emission from the planet at all orbital phases. The best fit brightness temperatures ranges from $1880\pm40$ K (phase $\phi = 0.1$) to $2930\pm10$ K on the dayside. We attribute the absence of water features at WFC3 wavelengths to (1) H$_2$O dissociation on the dayside and (2) additional near-IR opacity from H-, TiO/VO and FeH.}

>>>>

\begin{abstract}
	We present thermal phase curve measurements for the hot Jupiter WASP-103b observed with \emph{Hubble}/WFC3 and \emph{Spitzer}/IRAC. The dayside planet-to-star flux ratio is $0.151\pm0.015\%$, $0.446\pm0.038\%$, and $0.569\pm0.014\%$ in bandpasses centered at $1.4$, $3.6$, and $4.5\mu$m, respectively.  The phase curves have large amplitudes and negligible hotspot offsets, indicative of poor heat redistribution to the nightside.  We fit the phase variation with a range of climate maps and find that a spherical harmonics model generally provides the best fit. The phase-resolved spectra are consistent with blackbodies in the WFC3 bandpass, with brightness temperatures ranging from $1880\pm40$ K on the nightside to $2930 \pm 40$ K on the dayside. The dayside spectrum has a significantly higher brightness temperature in the \Spitzer\ bands, which we attribute to CO emission and a thermal inversion.  The inversion is not present on the nightside. The transmission spectrum is featureless, but consistent with the small features expected from cloud-free models. Using the dayside and nightside emission spectra, we retrieved the atmospheric composition with a 1D self-consistent modeling technique and find the atmosphere is moderately metal-enriched ($\mathrm{[M/H]} =  23^{+29}_{-13}\times$ solar) and the carbon-to-oxygen ratio is below 0.9 at $3\,\sigma$ confidence. In contrast to cooler hot Jupiters, we do not detect significant water absorption, which we attribute to partial H$_2$O dissociation and additional opacity from H- and trace amounts of hydrides/oxides.  We compare the phase curves to 3D general circulation models and find that atmospheric drag with a timescale of $10^4$\,s is the best match to the data, suggesting the presence of a magnetic field with strength $\sim1$ Gauss.  We also compare the WASP-103b spectra to brown dwarfs and young directly imaged companions and find these more massive and self-luminous objects have significantly larger water features, indicating that surface gravity and irradiation environment play an important role in shaping the spectra of hot Jupiters. These results highlight the 3D structure of exoplanet atmospheres and illustrate the importance of phase curve observations for understanding their complex chemistry and physics.
\end{abstract}

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/w103b_transmission_spectrum.pdf}
\caption{The transmission spectrum of WASP-103b, corrected for planet nightside emission at 1700 K (blue points) and uncorrected (red points). The dark gray line is the model transmission spectrum from the $\tau_\mathrm{drag4}$ GCM, with squares indicating the model binned over the \Spitzer\ bandpasses. Atmospheric scale height $H$ is shown on the right y-axis, where $H = 5.5\times10^6$ m (assuming a mean molecular weight of 2.3 atomic mass units, surface gravity $g = 15.9$ m/s$^2$, and a temperature $T = 2410$ K).}
\label{fig:tspec}
\end{figure}

@@

\subsection{Blackbody Fits}
\label{sec:bbfits}
A blackbody is the simplest model for the planet's thermal emission and provides a useful first evaluation of the atmospheric properties. To calculate the best fit blackbody model, we assumed a planet-to-star radius ratio of 0.1146 and used a NextGen stellar spectrum interpolated to an effective temperature of 6110 K \citep{allard12}.  We calculated the best fit with a least-squares fitting routine. To determine uncertainties on the planet brightness temperature, we performed an MCMC analysis with free parameters for the planet temperature and the stellar $T_\mathrm{eff}$. We used a Gaussian prior on $T_\mathrm{eff}$ of $6110 \pm 160$ K \citep{gillon14}.

The best-fit blackbodies are shown in Figures \ref{fig:spectra} and \ref{fig:dayside}. The model fits the data fairly well: it is consistent with the data at the $3\,\sigma$ level for 70\% of the orbital phases. However, the \Spitzer\ data have higher brightness temperatures than the WFC3 data at phase 0.5, and lower brightness temperatures at phases $0.8 - 0.9$. These differences suggest the presence of an emission feature on the dayside at \Spitzer\ wavelengths, which transitions to an absorption feature on the nightside, perhaps indicating changes in thermal structure with longitude in the atmosphere.

We also fit indepedent blackbody models to the \HST/WFC3 data and each \Spitzer\ channel. The resulting brightness temperatures and $1\,\sigma$ uncertainties are listed in Table\,\ref{table:teffs}. The WFC3 data agree well by a blackbody model at all orbital phases except phase 0.5 (consistent at better than $1.5\,\sigma$). The dayside has higher signal-to-noise than the other orbital phases, thanks to the two secondary eclipse observations from \cite{cartier17}. The more sophisticated grid-based retrieval (described in the next section) provides a better fit to the dayside.

@@

\begin{deluxetable}{LLLLL}
\tablecolumns{5}
\tablewidth{0pt}
\tablecaption{Phase-resolved Brightness Temperatures \label{table:teffs}}
\tablehead{
\colhead{Orbital Phase} & \colhead{$T_\mathrm{b}$} & \colhead{$T_\mathrm{b}$} & \colhead{$T_\mathrm{b}$} & \colhead{$\chi^2_\nu$\tablenotemark{1}} \\
\colhead{\,} & \colhead{WFC3} & \colhead{Ch. 1} & \colhead{Ch. 2} & \colhead{(9 dof)}}
\startdata
0.06 - 0.15 & 1883 \pm 41 & 1523 \pm 153 & 1589 \pm 105 & 0.7 \\
0.15 - 0.25 & 2208 \pm 33 & 2612 \pm 117 & 2299 \pm 100 & 0.9 \\
0.25 - 0.35 & 2587 \pm 37 & 2926 \pm 114 & 2592 \pm 96 & 1.5 \\
0.35 - 0.44 & 2831 \pm 39 & 2834 \pm 111 & 2976 \pm 93 & 1.5 \\
0.44 - 0.56 & 2933 \pm 41 & 2995 \pm 159 & 3154 \pm 99 & 2.8 \\
0.56 - 0.65 & 2811 \pm 39 & 2631 \pm 110 & 2891 \pm 97 & 2.0 \\
0.65 - 0.75 & 2572 \pm 36 & 2708 \pm 117 & 2421 \pm 100 & 1.3 \\
0.75 - 0.85 & 2263 \pm 33 & 1952 \pm 125 & 1939 \pm 99 & 1.3 \\
0.85 - 0.94 & 1987 \pm 37 & 1594 \pm 145 & 1288 \pm 118 & 0.6 \\
\enddata
\tablenotetext{1}{$\chi^2_\nu$ values are for the fits to the WFC3 data only.}
\end{deluxetable}

@@

\begin{deluxetable*}{llCCCC}
	\tablecolumns{6}
	\tablewidth{0pt}
	\tablecaption{Source Properties \label{table:sources}}
	\tablehead{
	\colhead{\,} & \colhead{Object} & \colhead{$T_\mathrm{eff}$ (K)} & \colhead{log\,$g$ (cgs)} & \colhead{H$_2$O A$_1$} & \colhead{H$_2$O A$_2$}}
		\startdata
		Hot Jupiter & W103b night & 1880 \pm 40 & 3.2 \pm 0.04 & 0.07 \pm 1.8\mathrm{e}-01 & -0.00 \pm 1.6\mathrm{e}-01 \\
		\, & W103b quadrature & 2400 \pm 40 & 3.2 \pm 0.04 & -0.14 \pm 7.8\mathrm{e}-02 & -0.01 \pm 6.7\mathrm{e}-02 \\
		\, & W103b dayside & 2930 \pm 40 & 3.2 \pm 0.04 & 0.04 \pm 1.4\mathrm{e}-02 & 0.15 \pm 1.2\mathrm{e}-02 \\
		Brown Dwarf & 2MASS J1320+0409 & 1880 \pm 70 & 5.19 \pm 0.16 & 0.21 \pm 6.3\mathrm{e}-04 & -0.06 \pm 5.0\mathrm{e}-04 \\
		\, & 2MASS J0428-2253 & 2430 \pm 80 & 5.22 \pm 0.09 & 0.16 \pm 1.2\mathrm{e}-04 & -0.03 \pm 1.0\mathrm{e}-04 \\
		\, & 2MASS J0003-2822 & 2890 \pm 80 & 5.18 \pm 0.04 & 0.26 \pm 1.3\mathrm{e}-04 & 0.10 \pm 1.1\mathrm{e}-04 \\
		Imaged Companion & CD-35 2722 & 1800 \pm 100 & 4.5 \pm 0.5 & - & 0.15 \pm 1.0\mathrm{e}-05 \\
		\, & USco 1610-1913B & 2400 \pm 150 & - & 0.27 \pm 2.0\mathrm{e}-04 & 0.19 \pm 2.2\mathrm{e}-04 \\
		\, & TWA 22A & 3000 \pm 100 & 4.5 \pm 0.5 & - & 0.29 \pm 2.0\mathrm{e}-05 \\
		\enddata
	\end{deluxetable*}

@@

	\item{The dayside planet-to-star flux is $0.151\pm0.015\%$, $0.446\pm0.38\%$, and $0.569\pm0.014\%$ in the WFC3 bandpass, \Spitzer\ $3.6$, and \Spitzer\ $4.5\mu$m, respectively.  The best fit blackbody to the WFC3 dayside spectrum has a brightness temperature of $2930 \pm 40$ K, making WASP-103b among the hottest exoplanets ever observed.}

@@

\item{We calculated phase-resolved spectra in ten orbital phase bins. The \HST/WFC3 spectra are consistent with  blackbody emission from the planet at all orbital phases. The best fit brightness temperatures ranges from $1880\pm40$ K (phase $\phi = 0.1$) to $2930\pm40$ K on the dayside. We attribute the absence of water features at WFC3 wavelengths to (1) H$_2$O dissociation on the dayside and (2) additional near-IR opacity from H-, TiO/VO and FeH.}
